
\beginhymn Conditor alme siderum.

\setgregorian1
\setclefsymbol1\gregorianFclef
\setclef15
\musicindent14mm
\raisesong3\Internote
\nosolesmescustos
\initiumgregorianum
%%%Creator alme siderum,
\musicinitial{Hymn.\\IV.}{C}%
\sgn {}O{n}\punctum L\egn
\sgn {d}{\'\i}{t}\punctum J\egn
\sgn {}or\engl{Bright builder of the heav'nly poles,}\punctum L\egn
\spatium
\sgn {}{a}{lm}\punctum N\egn
\sgn {}{e}{}\punctum N\egn
\spatium
\sgn {s}{\'\i}{d}\punctum a\egn
\sgn {}{e}{r}\punctum M\egn
\sgn {}u{m,}\punctum N\egn
\spatium
\divisiominima
\spatium
%%%AEterna lux cred\'entium,
\sgn {}{\AE}t\punctum N\egn
\sgn {}{\'e}{r}\punctum a\egn
\sgn na{}\punctum M\egn
\spatium
\sgn lux\punctum N\egn
\spatium
\sgn {cr}ed\engl{Eternal light of faithful souls,}\punctum M\egn
\sgn {}{\'e}n\punctum L\egn
\sgn {t}{i}{-}\punctum K\egn
\sgn {}u{m,}\punctum L\augmentum L\egn
\spatium
\divisiominor
\spatium
\custos N\lineaproxima
%%%J\'esu, Red\'emptor \'omnium,
\sgn {Chr}{i}s\punctum N\egn
\sgn {t}{e}{,}\punctum M\egn
\spatium
\sgn Red\punctum K\egn
\sgn {}{\'e}{m}\engl{Jesus, Redeemer of mankind,}\punctum L\egn
\sgn {pt}or\punctum M\egn
\spatium
\sgn {}{\'o}m\punctum L\egn
\sgn {n}{i}{-}\punctum K\egn
\sgn {}u{m,}\punctum J\egn
\spatium
\divisiominima
\spatium
%%%Int\'ende v\'otis s\'upplicum.
\sgn {}{E}{x}\punctum J\egn
\sgn {}{\'a}u\punctum L\egn
\sgn di{}\punctum M\egn
\spatium
\sgn {pr}{e}c\punctum N\egn
\sgn {}es\punctum M\egn
\spatium
\sgn {s}{\'u}p\engl{Our humble prayers vouchsafe to mind.}\punctum L\egn
\sgn {pl}{i}{c}\punctum K\egn
\sgn {}u{m.}\punctum L\augmentum L\egn
\spatium
\Finisgregoriana

\bigskip

\beginlyrics

\sidx{Creator of the starry skies}

\hbox{

\vtop{
\hsize42mm

%\rmlyrics

\verse2. Qui c\'ondolens int\'eritu\\
Mortis per\'\i re s\'\ae culum,\\
Salv\'asti mundum, l\'anguidum,\\
Donans reis rem\'edium.

\verse3. Verg\'ente mundi v\'espere,\\
Uti sponsus de th\'alamo,\\
Egr\'essus honest\'\i ssima\\
V\'{\i}rginis Matris cl\'ausula.

\verse4. Cujus fortis pot\'esti\ae \\
Genu curv\'antur \'omnia,\\
C{\ae}l\'estia, terr\'estria,\\
Nutu fat\'entur s\'ubdita.

\verse5. Te deprec\'amur h\'agie,\\
Vent\'ure judex s\'\ae culi,\\
Cons\'erva nos in t\'empore,\\
Hostis a telo p\'erfidi.

\verse6. Laus, honor, virtus, gl\'oria\\
Deo Patri, et F\'{\i}lio,\\
Sancto simul Par\'aclito,\\
In s{\ae}cul\'orum s\'{\ae}cula.

}

\hskip5mm

\vtop{
\hsize58mm

\it%lyrics

\verse2. Who, lest the fraud of hell's black king\\
Should all men to destruction bring,\\
Didst, by an act of gen'rous love,\\
The fainting world's physician prove.

\verse3. Thou, that Thou mightst our ransom pay\\
And wash the stains of sin away,\\
Didst from a Virgin's womb proceed\\
And to the Cross a Victim bleed.

\verse4. Thy glorious power, Thy saving name\\
No sooner any voice can frame,\\
But heaven and earth and hell agree\\
To honour them with trembling knee.

\verse5. Thee, Christ, who at the latter day\\
Shalt be our Judge, we humbly pray\\
Such arms of heav'nly grace to send\\
As may Thy Church from foes defend.

\verse6. Be glory given and honour done\\
To God the Father, and the Son\\
And to the Holy Ghost on high,\\
From age to age eternally.

}

}


\endlyrics

\medskip

\source Ambrosian, 7th century

\source Translation from the Primer {\smold 1685} and the Evening Office \smold 1710

