
\beginhymn Corda pia inflammantur.

\setgregorian1
\setclef14
\musicindent14mm
\raisesong3\Internote
\nosolesmescustos
\initiumgregorianum
%%%Ave verum
\musicinitial{VI}{C}%
\sgn {}{O}r\punctum M\egn
\sgn da{}\punctum N\egn
\spatium
\sgn {p}{i}{-}\punctum a\egn
\sgn {}{a}{}\engl{Godly hearts are inflamed}\punctum a\egn
\spatium
\sgn {}{i}{nfl}\bmolle b\punctum b\egn
\sgn {}{a}{m}\punctum b\egn
\sgn {m}{\'a}{nt}\punctum a\egn
\sgn {}{u}{r,}\punctum a\augmentum b\egn
\spatium
\divisiominor
\spatium
\sgn {D}{u}{m}\punctum c\egn
\spatium
\sgn {Fr}{a}{n}\bmolle b\punctum b\egn
\sgn {c}{\'\i}{s}\engl{while  the tokens}\punctum b\egn
\sgn {c}{i}{}\punctum a\egn
\spatium
\sgn {c}{e}{-}\punctum a\egn
\sgn {l}{e}{-}\bmolle b\punctum b\egn
\custos N\lineaproxima
\sgn {br}{\'a}{n}\punctum N\egn
\sgn {t}{u}{r}\punctum N\augmentum N\egn
\spatium
\divisiominima
\spatium
\sgn {St}{\'\i}{g}\punctum M\egn
\sgn {m}{a}{-}\punctum N\egn
\sgn {t}{u}{m}\engl{of Francis' Stigmata are celebrated.}\punctum a\egn
\spatium
\sgn {}{i}{n}\bmolle b\punctum b\egn
\sgn {s}{\'\i}{g}\punctum N\egn
\sgn {n}{i}{-}\punctum M\egn
\sgn {}{a}{}\punctum M\augmentum N\egn
\Finisgregoriana

\bigskip


\beginlyrics

%\it Et a Choro repetitur: \rm

%Corda pia inflamm\'antur,\\
%Dum Franc{\'\i}sci celebr\'antur\\
%St{\'\i}gmatum ins{\'\i}gnia.\\
%\quad \it Cantores prosequuntur: \rm

\hbox{

\vtop{
\hsize47mm

\verse1. Absit nobis glori{\'a}ri,\\
Nisi\lyrlink in Cruce salut\'ari\\
Franc{\'\i}sci vest{\'\i}gio.%\\
%\quad \it Chorus: \rm Corda pia.

\verse2. Nam in monte vir dev\'otus,\\
Vigil, nudus, ardens totus\\
Crebra dat susp{\'\i}ria.%\\
%\quad \it Chorus: \rm Corda pia.

\verse3. Solus ergo cl\'ausus orans,\\
Super gestis Crucis plorans,\\
M\oe r\'ore conf{\'\i}citur.%\\
%\quad \it Chorus: \rm Corda pia.

\verse4. Cernit servus Redempt\'orem,\\
S\ae cul\'orum Imperat\'orem,\\
Passum impass{\'\i}bilem.%\\
%\quad \it Chorus: \rm Corda pia.

\verse5. Cor Franc{\'\i}sci transform\'atur,\\
Corpus vero mox orn\'atur\\
Mir\'andis stigm\'atibus.%\\
%\quad \it Chorus: \rm Corda pia.


\verse6. Crucif{\'\i}xi ergo Christi\\
Mors et vita fuit isti\\
Jugis medit\'atio.%\\
%\quad \it Chorus: \rm Corda pia.

\verse7. Cujus cordis vim fev\'oris\\
Ost\'endit per membra foris\\
St{\'\i}gmatum impr\'essio.%\\
%\quad \it Chorus: \rm Corda pia.

}
\hskip5mm
\vtop{
\hsize53mm

\it

\verse1. Far be it from us to glory, 
save in the healing Cross, 
in the footsteps of Francis.

\verse2. For the devoted man on the mountain,
keeping watch, naked, wholly burning,
frequently sighs.

\verse3. Therefore, alone, praying with tears
meditating on the history of the Cross,
he is transfixed with sorrow.

\verse4. The servant sees his Redeemer,
the commander of ages,
the unassailable assailed.

\verse5. Francis' heart is transformed;
soon indeed his body is adorned with 
the marvellous Stigmata.

\verse6. Therefore the life and death
of Christ crucified was his
constant meditation.

\verse7. The impression of the Stigmata,
shows the force of his heart's fervour
outwardly through his limbs.

}
}

\hbox{

\vtop{

\hsize47mm


\verse8. Crucif{\'\i}xe, singul\'ari\\
Modo Cruci conform\'ari\\
Mente fac et h\'abitu.%\\
%\quad \it Chorus: \rm Corda pia.

\verse9. Fac ut nos in regno lucis\\
Perfru\'amur fructu Crucis,\\
Quo l\ae t\'emur c\'\ae litus.%\\
%\quad \it Chorus: \rm Corda pia.

\verse10. Ad quem venit Rex e c\ae lo,\\
Affix\'usque Crucis telo\\
Asp\'ectu pac{\'\i}fico.%\\
%\quad \it Chorus: \rm Corda pia.

\verse11. Collaud\'etur Crucif{\'\i}xus,\\
Franc{\'\i}scus prorsus inn{\'\i}xus\\
Super mundi f\'\oe dera.%\\
%\it Chorus: \rm Corda pia.

}
\hskip5mm
\vtop{
\hsize53mm
\it

\verse8. O Crucifed One, grant that we may be conformed in our 
mind and actions 
to the pattern of the Cross.

\verse9. Make us benefit from the fruit of the Cross
in the kingdom of light
where we may rejoice with the heavenly one.

\verse10. To whom the King of Heaven
came, fastened to the cross with serene countenance.

\verse11. Let the Crucified One be praised
with Francis who is resting
above the plottings of the world.

}

}


\endlyrics

\bigskip

%\source With thanks to Br.~Seraphim.

%Also, aim to pursue a translation

%\endhymn

