%Jesu Dulcis Memoria

\beginhymn Jesu dulcis memoria.

\setgregorian1
\musicindent14mm
\raisesong3\Internote
\nosolesmescustos
\setclef14
\initiumgregorianum
\musicinitial{Hymn\\I.}{J}%
\sgn {}Es\punctum a\egn
\sgn {}u{}\punctum a\egn
\spatium
\sgn d{u}l\punctum a\egn
\sgn cis\engl{\idx{Jesus, the very thought of Thee}}\punctum a\egn
\spatium
\sgn mem\punctum K\egn
\sgn {}{\'o}r\punctum M\egn
\sgn {}i-\punctum L\egn
\sgn {}a{,}\punctum K\egn
\spatium
\divisiominima
\spatium
\sgn Da{ns}\punctum a\egn
\spatium
\sgn v{e}r\punctum a\egn
\sgn {}a{}\punctum b\egn
\spatium
\sgn c{o}r\punctum N\egn
\sgn dis\engl{With sweetness fills my breast,}\punctum a\egn
\spatium
\sgn g{\'a}{ud}\punctum c\egn
\sgn {}i-\punctum b\egn
\sgn {}a{}\punctum a\augmentum b\egn
\spatium
\divisiominor
\spatium
\custos a
\lineaproxima
\sgn Sed\punctum a\egn
\spatium
\sgn sup\punctum b\egn
\sgn {}er\punctum c\egn
\spatium
\sgn mel\engl{But sweeter far Thy face to see}\punctum d\egn
\spatium
\sgn {}et\punctum c\egn
\spatium
\sgn {}{\'o}m\punctum a\egn
\sgn ni-\punctum c\egn
\sgn {}a{,}\climacus baN\egn
\spatium
\divisiominor
\spatium
\sgn {}Ej\punctum a\egn
\sgn {}us\punctum a\egn
\spatium
\sgn d{u}l\punctum b\egn
\sgn cis\punctum N\egn
\spatium
\sgn {pr}{\ae}s\engl{And in Thy presence rest.}\punctum L\egn
\sgn {}{\'e}{n-}\punctum N\egn
\sgn ti-\punctum a\egn
\sgn {}a{}\punctum a\augmentum b\egn
\spatium
\Finisgregoriana

\bigskip

\beginlyrics

\hbox{

\vtop{
\hsize41mm

\verse2. Nil c\'anitur su\'avius,\\
Nil aud\'\i tur juc\'undius,\\
Nil cogit\'atur d\'ulcius,\\
Quam Jesus Dei F\'\i lius.

\verse3. Jesu spes p\ae nit\'entibus,\\
Quam pius es pet\'entibus!\\
Quam bonus te qu\ae r\'en\-ti\-bus!\\
Sed quid inveni\'entibus?

\verse4. Nec lingua valet d\'\i cere,\\
Nec l\'\i ttera expr\'\i mere:\\
Exp\'ertus potest cr\'edere,\\
Quid sit Jesum dil\'\i gere.

\verse5. Sis Jesu nostrum g\'audium,\\
Qui es fut\'urus pr\'\ae mium:\\
Sit nostra in te gl\'oria,\\
Per cuncta semper s\'\ae cula.

Amen.

}

\hskip7mm

\vtop{
\hsize57mm

\it

\verse2. No voice can sing, no heart can frame,\\
Nor can the memory find,\\
A sweeter sound than Thy blest Name,\\
O Saviour of mankind.

\verse3. O hope of every contrite heart,\\
O joy of all the meek,\\
To those who fall, how kind Thou art!\\
How good to those who seek.


\verse4. But what to those that find? Ah! this\\
No tongue nor pen can show;\\
The love of Jesus, what it is,\\
None but His loved ones know.


\verse5. Jesu, our only joy be Thou,\\
As Thou our prize wilt be,\\
Jesu, be Thou our glory now,\\
And through eternity.

Amen.

}

}

\endlyrics

%\bigskip


\bigskip

\source Attributed to St.~Bernard of Clairvaux, {\smold1091}--\smold1153

\source Translated by Fr.~Edward Caswall, {\smold1814}--\smold78

